---
title: 'Lab 1'
author: 'Angelo Dagonel'
date: '8/31/2017'
output: github_document
---

```{r include=FALSE}

### Loading necessary packages

library(dplyr)
library(printr)
library(tidyr)
library(knitr)
library(scales)
library(ggplot2)

# Get rid of warning messages
options(warn=-1)
```

---

#R as a calculator

**Calculate Leia's total expenses**

cell phone        $80 
transportation    $20
groceries         $527
gym               $10
rent              $1500
other             $83

*Insert chunk:*       `Ctrl+Alt+I / Cmd+Option+I`

*Run current chunk:*  `Ctrl+Shift+Enter / Cmd+Shift+Enter?`


Basic addition
```{r}
#Total expenses
80 + 20 + 527 + 10 + 1500 + 83
```

Creating objects and adding with operators `<-`
```{r}
#Creating objects with values using <-
phone <- 80
transportation <- 20
groceries <- 527
gym <- 10
rent <- 1500
other <- 83

#Example: Inspect value of cell phone expenses
phone

#Total expenses via Sum of objects
total <- phone + transportation + groceries + gym + rent + other
total

#Total semester expenses (5 months)
semtot <- total*3
semtot

#Total school year (10 months, or 2 semesters)
yeartot <- semtot*2
yeartot
```

#Functions

Absolute values
```{r}
abs(10)
abs(-4)
```

Square roots
```{r}
sqrt(9)
sqrt(16)
```

Natural logarithms
```{r}
log(2)
```

#Vectors

Create vector of Leia's expenses using combine function `c()`

```{r}
#Vector expenses
expenses <- c(phone, transportation, groceries, gym, rent, other)
names(expenses) <- c('Cell phone', 'Transportation', 'Groceries', 'Gym', 'Rent', 'Other')
expenses

#Barplot of expenses
barplot(expenses)

#Add names to values to make data frame
expenses_names <- c('Cell phone', 'Transportation', 'Groceries', 'Gym', 'Rent', 'Other')
expenses_table <- data.frame(expenses)
expenses_table
```

Sort vector in decreasing order, and add values to top of bar graph

```{r}
expenses_sort <- sort(expenses, decreasing=TRUE)

expenseplot <- barplot(sort(expenses, decreasing=TRUE))
expenseplot

text(x=expenseplot,
     y=expenses_sort,
     label=expenses_sort,
     pos=3,
     cex=0.8)
```


#Pythagoras formula

```{r}
#Create 'a' and 'b' for Pythagoras
a = 3
b = 4

c = sqrt(a^2 + b^2)
c
```

#Binomial formula

```{r}
#Compute number of combinations
choose(n=5, k=2)

#Factorial
factorial(4)
```

**Fair coin toss**

Fair coin is tossed *n=5 times*--What's the *probability* of getting  *exactly 2 heads*?

```{r}
#Assign objects for use in calculation
n = 5
p = 0.5
k = 2

#Combination calculation with `factorial`
comb_factorial = factorial(n)/(factorial(k)*factorial(n-k))
comb_factorial

#Combination calculation with Choose
comb_choose = choose(n, k)
comb_choose

#Calculate probability including Factorial
(factorial(n)/(factorial(k)*factorial(n-k)))*(p^k)*((1-p)^(n-k))
#or
prob1 = comb_factorial*(p^k)*((1-p)^(n-k))
prob1

#Calculate probability including Choose
choose(n, k)*(p^k)*((1-p)^(n-k))
#or
prob2 = comb_factorial*(p^k)*((1-p)^(n-k))
prob2
```

**Fair die roll**

Fair die is rolled *n=10 times*--What's the probability of rolling *exactly k=3 6s*?

```{r}
#Assignment of objects
n = 10
p = 1/6
k = 3

#Combination via `choose`
comb_die1 = choose(n,k)
comb_die1

#Probability including Choose
prob_die1 = comb_die1*(p^k)*((1-p)^(n-k))
prob_die1
```

**`dbinom()` function**

`dbinom(x, size, prob, log = FALSE)`
or
`dbinom(k, n, prob)`

```
x, q    vector of quantiles.
p	      vector of probabilities.
n	      number of observations (If length(n)>1, length = # required)
size    number of trials (zero or more).
prob	  probability of success on each trial.
log/.p  logical; if TRUE, probabilities p are given as log(p).
```

2 heads in 5 coin tosses
```{r}
dbinom(2, 5, 0.5)
```

3 sixes in 3 rolls of a die
```{r}
dbinom(3, 3, 1/6)
```

2 heads in 5 coin tosses, biased coin (35%)
```{r}
dbinom(2, 5, .35)
```
